#Packages
```{r}
library(dplyr)
library(phyloseq)
library(vegan)
library(openxlsx)
library(tidyverse)
library(emmeans)
library(car)
library(broom)
library(ggsignif)
library(glue)
library(tibble)
library(ggfortify)
library(patchwork)
library(VennDiagram)
library(ggvenn)
library(readxl)
library("RColorBrewer")
library("ggplot2")
library("gplots")
library("ggplot2")
library("pheatmap")
library("RColorBrewer")
library("tibble")
library("tidyverse")
library("dplyr")
library("openxlsx")
library("AID")
library("plotrix")
library("UpSetR")

# Install BiocManager if it's not already installed
if (!requireNamespace("BiocManager", quietly = TRUE)) {
    install.packages("BiocManager")
}

BiocManager::install("mixOmics")
library(mixOmics)
```
#Load metab
```{r}

#For rumen concentration load corresponding data for metab
library(readxl)
metab <- read_excel("metab_rumen_conc.xlsx")
View(metab)
```
#Long version of metab
```{r}
metablong <- metab %>%
  pivot_longer(
    cols = starts_with("m_"),  # Select columns starting with "m_"
    names_to = "Metabolite",          # Name of the new column for metabolite
    values_to = "Value"          # Name of the new column for values
  )
view(metablong)
str(metablong)

# Remove "m_" from the 'Metabolite' column
metablong$Metabolite <- gsub("^m_", "", metablong$Metabolite)
view(metablong)
```
#Load metabolite classification
```{r}
library(readxl)
class <- read_excel("class.xlsx")
View(class)

# Joining metablong with class based on Metabolite column
metablong <- metablong %>%
  left_join(class, by = "Metabolite")

# View the resulting data frame
view(metablong)
```
#Averages of every metabolite
```{r}
# Calculate the average value for each combination of treatment and Metabolite
average_treatment <- metablong %>%
  group_by(Treatment,Metabolite) %>%
  summarize(AverageValue = mean(Value, na.rm = TRUE)) %>%
  ungroup()

# View the resulting data frame
view(average_treatment)
write.table(average_treatment, file = "average_treatment.txt")

# Calculate the average value for each combination of period and Metabolite
average_period <- metablong %>%
  group_by(Period,Metabolite) %>%
  summarize(AverageValue = mean(Value, na.rm = TRUE)) %>%
  ungroup()

# View the resulting data frame
view(average_period)
write.table(average_period, file = "average_period.txt")

# Calculate the average value for each combination of TP and Metabolite
average_tp <- metablong %>%
  group_by(TP,Metabolite) %>%
  summarize(AverageValue = mean(Value, na.rm = TRUE)) %>%
  ungroup()

# View the resulting data frame
view(average_tp)
write.table(average_tp, file = "average_tp.txt")

rm(average_treatment, average_period, average_tp)
```
#SEM of every metabolite
```{r}
# Calculate the standard error for each combination of Metabolite
se_treat <- metablong %>%
  group_by(Metabolite) %>%
  summarize(
    Count = n(),
    StandardError = sd(Value, na.rm = TRUE) / sqrt(n())
  ) %>%
  ungroup()

# View the resulting data frame
view(se_treat)
write.table(se_treat, file = "se_treat.txt")

rm(se_treat)

```
#Shapiro and Levene test for every metabolite
```{r}
library(purrr)

# Assuming your data frame is named metablong and has columns: Metabolite and Value

# Perform Shapiro-Wilk test for each metabolite
shapiro_results <- metablong %>%
  group_by(Metabolite) %>%
  summarize(
    W_statistic = tryCatch(shapiro.test(Value)$statistic, error = function(e) NA),
    p_value = tryCatch(shapiro.test(Value)$p.value, error = function(e) NA)
  ) %>%
  ungroup()

# View the resulting data frame
view(shapiro_results)
write.table(shapiro_results, file = "shapiro_results.txt")

# Perform Levene's test for each metabolite
levene_results <- metablong %>%
  group_by(Metabolite) %>%
  summarize(
    Test_Statistic = tryCatch(leveneTest(Value ~ Treatment)$`F value`[1], error = function(e) NA),
    p_value = tryCatch(leveneTest(Value ~ Treatment)$`Pr(>F)`[1], error = function(e) NA)
  ) %>%
  ungroup()

# View the resulting data frame
print(levene_results)
write.table(levene_results, file = "levene_results.txt")

rm(shapiro_results, levene_results)
```
#Testing out the shapiro and levene of 3 metabolites
```{r}
shapiro.test(metab$`m_3-Hydroxyphenylacetate`)
leveneTest(`m_3-Hydroxyphenylacetate`~Treatment, data = metab)
kruskal.test(`m_3-Hydroxyphenylacetate`~Treatment, data = metab )
kruskal.test(`m_3-Hydroxyphenylacetate`~Period, data = metab )
kruskal.test(`m_3-Hydroxyphenylacetate`~TP, data = metab )

kruskal.test(m_Succinate~Treatment, data = metab )
kruskal.test(m_Succinate~Period, data = metab )
kruskal.test(m_Succinate~TP, data = metab )

#Works for Levene but not for ANOVA - need to ANOVA

shapiro.test(metab$m_Isobutyrate)
leveneTest(m_Isobutyrate~Treatment, data = metab)
IB<- aov(m_Isobutyrate ~ Treatment*Period, data = metab)
summary(IB)

shapiro.test(metab$m_Tyrosine)
leveneTest(m_Tyrosine ~ Treatment, data = metab)
Tyr<- aov(m_Tyrosine ~ Treatment*Period, data = metab)
summary(Tyr)
```
#How to select Metabolites that have a <0.05 shapiro and Perform Kruskal
```{r}
# Filter metabolites with p-value < 0.05
significant_metabolites <- shapiro_results %>%
  filter(p_value < 0.05) %>%
  pull(Metabolite)  # Extract the list of significant metabolites

#Perform Kruskal-Wallis test for each significant metabolite
kruskal_results <- metablong %>%
  filter(Metabolite %in% significant_metabolites) %>%
  group_by(Metabolite) %>%
  summarize(
    KW_treatment = tryCatch(kruskal.test(Value ~ Treatment)$p.value, error = function(e) NA),
    KW_period = tryCatch(kruskal.test(Value ~ Period)$p.value, error = function(e) NA),
    KW_TP = tryCatch(kruskal.test(Value ~ TP)$p.value, error = function(e) NA)
  ) %>%
  ungroup()

# View the resulting data frame
print(kruskal_results)
write.table(kruskal_results, file = "kruskal_results.txt")

rm(significant_metabolites, kruskal_results)
```
#How to select Metabolites that have a >0.05 shapiro and Perform Anova
```{r}
# Filter metabolites with p-value < 0.05
insignificant_metabolites <- shapiro_results %>%
  filter(p_value > 0.05) %>%
  pull(Metabolite)  # Extract the list of significant metabolites

# Step 2: Perform ANOVA for each normal metabolite
anova_results <- metablong %>%
  filter(Metabolite %in% insignificant_metabolites) %>%
  group_by(Metabolite) %>%
  summarize(
    ANOVA_Treatment = tryCatch(
      summary(aov(Value ~ Treatment))[[1]]["Treatment", "Pr(>F)"],
      error = function(e) NA
    ),
    ANOVA_Period = tryCatch(
      summary(aov(Value ~ Period))[[1]]["Period", "Pr(>F)"],
      error = function(e) NA
    ),
    ANOVA_TP = tryCatch(
      summary(aov(Value ~ TP))[[1]]["TP", "Pr(>F)"],
      error = function(e) NA
    )
  ) %>%
  ungroup()

# View the resulting data frame
print(anova_results)
write.table(anova_results, file = "anova_results.txt")

#Need to manually do ANOVA

```
#Manual ANOVA
```{r}
shapiro.test(metab$`m_3-Phenylpropionate`)
leveneTest(`m_3-Phenylpropionate`~Treatment, data = metab)
pp3<- aov(`m_3-Phenylpropionate` ~ Treatment*Period, data = metab)
summary(pp3)
rm(pp3)

shapiro.test(metab$m_Acetate)
leveneTest(m_Acetate~Treatment, data = metab)
a<- aov(m_Acetate ~ Treatment*Period, data = metab)
summary(a)
rm(a)

shapiro.test(metab$m_Aspartate)
leveneTest(m_Aspartate~Treatment, data = metab)
asp<- aov(m_Aspartate ~ Treatment*Period, data = metab)
summary(asp)
rm(asp)

shapiro.test(metab$m_Benzoate)
leveneTest(m_Benzoate~Treatment, data = metab)
B<- aov(m_Benzoate ~ Treatment*Period, data = metab)
summary(B)
rm(B)

shapiro.test(metab$m_Butyrate)
leveneTest(m_Butyrate~Treatment, data = metab)
By<- aov(m_Butyrate ~ Treatment*Period, data = metab)
summary(By)
rm(By)

shapiro.test(metab$m_Glutamate)
leveneTest(m_Glutamate~Treatment, data = metab)
g<- aov(m_Glutamate ~ Treatment*Period, data = metab)
summary(g)
rm(g)

shapiro.test(metab$m_Isobutyrate)
leveneTest(m_Isobutyrate~Treatment, data = metab)
IB<- aov(m_Isobutyrate ~ Treatment*Period, data = metab)
summary(IB)
rm(IB)

shapiro.test(metab$m_Lactose)
leveneTest(m_Lactose~Treatment, data = metab)
l<- aov(m_Lactose ~ Treatment*Period, data = metab)
summary(l)
rm(l)

shapiro.test(metab$`m_N-Phenylacetylglycine`)
leveneTest(`m_N-Phenylacetylglycine`~Treatment, data = metab)
nP<- aov(`m_N-Phenylacetylglycine` ~ Treatment*Period, data = metab)
summary(nP)
rm(nP)

shapiro.test(metab$m_Propionate)
leveneTest(m_Propionate~Treatment, data = metab)
pp<- aov(m_Propionate ~ Treatment*Period, data = metab)
summary(pp)
rm(pp)

shapiro.test(metab$m_Tyrosine)
leveneTest(m_Tyrosine ~ Treatment, data = metab)
Tyr<- aov(m_Tyrosine ~ Treatment*Period, data = metab)
summary(Tyr)
rm(Tyr)
```
#Heatmap
```{r}
library(dplyr)
library(tidyr)
library(pheatmap)

# Prepare the data for the heatmap
heatmap_data <- metablong %>%
  group_by(Metabolite, Treatment) %>%
  summarize(Average_Value = mean(Value, na.rm = TRUE)) %>%
  pivot_wider(names_from = Treatment, values_from = Average_Value, values_fill = list(Average_Value = 0))

# Convert to matrix for heatmap
heatmap_matrix <- as.matrix(heatmap_data %>% dplyr::select(-Metabolite) %>% column_to_rownames("Metabolite"))

# Generate the heatmap
pheatmap(
  heatmap_matrix,
  cluster_rows = TRUE,    # Cluster metabolites (rows)
  cluster_cols = TRUE,    # Cluster treatments (columns)
  scale = "row",          # Scale values across rows (metabolites)
  color = colorRampPalette(c("blue", "white", "red"))(50),  # Color gradient from blue to red
  main = "Heatmap of Metabolite Functionality by Treatment",
  filename = "heatmap_treatment.png"
)

```
#Heatmap with metabolism
```{r}
set.seed(123)

library(pheatmap)
library(dplyr)
library(tidyr)

# Ensure all values are numeric and handle non-numeric data
metablong <- metablong %>%
  mutate(Value = as.numeric(Value)) %>%
  filter(!is.na(Value))

# Ensure there are no conflicts with the summarize function
heatmap_data <- metablong %>%
  dplyr::group_by(Metabolite, Treatment) %>%
  dplyr::summarize(Average_Value = mean(Value, na.rm = TRUE), .groups = 'drop') %>%
  tidyr::pivot_wider(names_from = Treatment, values_from = Average_Value, values_fill = list(Average_Value = 0))

# Include MetabolismGroup information
heatmap_data <- heatmap_data %>%
  left_join(metablong %>% dplyr::select(Metabolite, Metabolism) %>% distinct(), by = "Metabolite")

# Convert to matrix for heatmap
heatmap_matrix <- as.matrix(heatmap_data %>% dplyr::select(-Metabolite, -Metabolism))
rownames(heatmap_matrix) <- heatmap_data$Metabolite

# Create a data frame for the row annotations
annotation_df <- data.frame(Metabolism = heatmap_data$Metabolism)
rownames(annotation_df) <- heatmap_data$Metabolite

# Define colors for metabolism groups if needed
unique_groups <- unique(heatmap_data$Metabolism)
annotation_colors <- list(Metabolism = setNames(RColorBrewer::brewer.pal(length(unique_groups), "Set3"), unique_groups))

# Reorder columns of heatmap_matrix to SIL, BG, TN
heatmap_matrix <- heatmap_matrix[, c("SIL", "BG", "TN")]

# Generate the heatmap with annotations
heatmap_grob<- pheatmap(
  heatmap_matrix,
  cluster_rows = TRUE,    # Cluster metabolites (rows)
  cluster_cols = FALSE,    # Cluster treatments (columns)
  scale = "row",          # Scale values across rows (metabolites)
  color = colorRampPalette(c("blue", "white", "red"))(50),  # Color gradient from blue to red,
  annotation_row = annotation_df,   # Row annotations for metabolism group
  annotation_colors = annotation_colors, # Colors for metabolism group annotations
  clustering_method = "average",
  clustering_distance_rows = "euclidean",
    treeheight_row = 10, # Adjust the height of the row dendrogram
  treeheight_col = 10,  # Adjust the height of the column dendrogram
   cellheight = 10,    # Increase to make cells taller
  cellwidth = 50,     # Increase to make cells wider
   )$gtable

heatmap_plot<- ggdraw(heatmap_grob)
heatmap_plot


```
#Load CSV for Metabolomic Exploratory analysis
```{r}
#Rumen concentration Metabolomics
library(readr)
rumen <- read_csv("rumen_metabolomics_conc.csv")
View(rumen)

```
#MIXOMICS PCA, PLSDA and OPLSDA
```{r}
library(mixOmics)
library(grid)
library(cowplot)
library(patchwork)
install.packages("magick")
library(magick)


#Removing meta data columns from rumen
rumen_data <- rumen[, -c(2:4)]

#Assigning Treatment for PCA
treatment <- rumen$Treatment
treatment <- rumen[[2]]
# Assuming you've extracted the 'Treatment' column correctly
treatment <- as.factor(treatment)


# Define colors for the legend
colors <- c("BG" = "darkgreen","SIL" = "blue",  "TN" = "grey")

# Convert the data frame to a matrix
rumen_matrix <- as.matrix(rumen_data)

#PCA results
pca_result <- pca(rumen_matrix, ncomp = 2)  # ncomp specifies the number of components to calculate
# Plot PCA
png("PCA_plot.png", width = 1800, height = 1200, res = 300)
plotIndiv(pca_result, group = treatment, legend = FALSE, title = "(a)",
          ellipse = TRUE, col.per.group = colors, pch = 19, legend.title = "Feed   Treatment")
dev.off()
pca_plot_grid<- ggdraw() + draw_image("PCA_plot.png")
pca_plot_grid

#PCA combined plot
pca_plot<- plotIndiv(pca_result, group = treatment, legend = TRUE, title = "(a)",ellipse = TRUE, col.per.group = colors, pch = 19, legend.title = "Feed   Treatment")
pca_plot


# PLS-DA
plsda_result <- plsda(rumen_matrix, treatment, ncomp = 2)
png("PLSDA_plot.png", width = 1800, height = 1200, res = 300)
plotIndiv(plsda_result, group = treatment, legend = FALSE, title = "(b)",
          ellipse = TRUE,col.per.group = colors, pch = 19, legend.title = "Feed   Treatment")
dev.off()
plsda_plot_grid<- ggdraw() + draw_image("PLSDA_plot.png")
plsda_plot_grid

#PLSDA combined plot
plsda_plot<- plotIndiv(plsda_result, group = treatment, legend = TRUE, title = "(b)",
          ellipse = TRUE,col.per.group = colors, pch = 19, legend.title = "Feed   Treatment")
plsda_plot

#Save combined plot
# Combine PCA and PLS-DA with a shared legend
upper_row<- (pca_plot_grid + plsda_plot_grid) + plot_layout(guides = "collect")
upper_row

# OPLS-DA (using splsda for implementation)
oplsda_result <- splsda(rumen_matrix, treatment, ncomp = 2)
png("OPLSDA_plot.png", width = 1800, height = 1200, res = 300)
plotIndiv(oplsda_result, group = treatment, legend = TRUE, title = "OPLS-DA Plot",
          ellipse = TRUE, col.per.group = colors, pch = 19, legend.title = "Feed   Treatment")
dev.off()

rm(pca_result, plsda_result, oplsda_result, rumen_data, rumen_matrix,
   treatment, colors)
```
#Combine PCA, PLSDA and Metabolites
```{r}
# Combine the upper row with the heatmap below
combined_plot <- (upper_row / heatmap_plot) + plot_layout(nrow = 2)
combined_plot

# Save the combined plot
ggsave("Metab_combined_plot.png", plot = combined_plot, width = 17, height = 24, units = "cm", dpi = 600)
```

#Correlation of metabolites
```{r}
#Test parametric / non parametric

shapiro.test(rumen$`3-Hydroxyphenylacetate`)
shapiro.test(rumen$Gentisate)

#therefore use spearman correlation

#Remove column 1-4 of rumen data
rumen_corr<- rumen[,-c(1:4)]

# Compute the Pearson correlation matrix using pairwise complete observations
correlation_matrix <- cor(rumen_corr, use = "pairwise.complete.obs", method = "spearman")

#Round data to two digits
correlation_matrix<- round(correlation_matrix, digits = 2)

# improved correlation matrix
library(corrplot)
png("Rumen_metab_Spearman_Corr.png", width = 800, height = 600)
corrplot((correlation_matrix),
             method = "circle",
             type = "upper", 
             tl.cex = 1,        # Text label size
             number.cex = 1.5,    # Correlation number size
              mar = c(0,0, 0, 0)  # Margin adjustment
         )
dev.off()

rm( rumen_corr, correlation_matrix)
```
#Significant corr plot
```{r}
library(Hmisc)

#Remove column 1-4 of rumen data
rumen_corr<- rumen[,-c(1:4)]

# Calculate Spearman correlation matrix and p-values
correlation_result <- rcorr(as.matrix(rumen_corr), type = "spearman")

# Extract correlation matrix and p-value matrix
correlation_matrix <- correlation_result$r
p_value_matrix <- correlation_result$P

# Set significance level
significance_level <- 0.05

# Create a mask for significant correlations
significant_mask <- p_value_matrix < significance_level

# Set non-significant correlations to NA
correlation_matrix[!significant_mask] <- NA

# Plot the correlation matrix with significant correlations only
png("Rumen_metab_Spearman_Corr_SIG.png", width = 800, height = 600)
corrplot(correlation_matrix,
         method = "circle",
         type = "upper",
         tl.cex = 1,        # Text label size
         number.cex = 1.5,  # Correlation number size
         mar = c(0, 0, 0, 0),  # Margin adjustment
         na.label = " ",   # Remove non-significant values
         na.label.col = "white"  # Color for non-significant values
)
dev.off()

rm(correlation_matrix, correlation_result, rumen_corr, significant_mask, 
   significance_level, p_value_matrix)
```




